---
title: "DataSetupOCExample"
author: "Jon Ahlinder"
date: "2023-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(nadiv)
library(rstan)
require(BGLR)
library(Matrix)
options(mc.cores = parallel::detectCores()) # For execution on a local, multicore CPU with excess RAM
rstan_options(auto_write = TRUE) # To avoid recompilation of unchanged Stan programs
library(brms)
library(corpcor) # load corpcor library to fix positive definiteness
library(asreml)
library(asremlPlus)
library(stringr)
#library(INLA)
library(tibble)
library(AGHmatrix)
library(pheatmap)
path <- "/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/"
```

## Notes

NB! Use the same pedigree (and relationship matrix) in both Opsel and OSQP

## Read Phenotypic data

```{r}
S22F812A2 <- read.table(file = paste(path,"DMS_stems_S22F812A2.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S22F812A2:\n'))
print(colnames(S22F812A2)[grep('Adj_',colnames(S22F812A2))])
S23F812A1 <- read.table(file = paste(path,"DMS_stems_S23F812A1.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F812A1:\n'))
print(colnames(S23F812A1)[grep('Adj_',colnames(S23F812A1))])
S23F832392 <- read.table(file = paste(path,"DMS_stems_S23F832392.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F832392:\n'))
print(colnames(S23F832392)[grep('Adj_',colnames(S23F832392))])
S23F832393 <- read.table(file = paste(path,"DMS_stems_S23F832393.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F832393:\n'))
print(colnames(S23F832393)[grep('Adj_',colnames(S23F832393))])
S23F832394 <- read.table(file = paste(path,"DMS_stems_S23F832394.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F832394:\n'))
print(colnames(S23F832394)[grep('Adj_',colnames(S23F832394))])
S23F832395 <- read.table(file = paste(path,"DMS_stems_S23F832395.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F832395:\n'))
print(colnames(S23F832395)[grep('Adj_',colnames(S23F832395))])
S23F8820476 <- read.table(file = paste(path,"DMS_stems_S23F8820476.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F8820476:\n'))
print(colnames(S23F8820476)[grep('Adj_',colnames(S23F8820476))])
S23F8820477 <- read.table(file = paste(path,"DMS_stems_S23F8820477.txt",sep=""),sep=",",header=T) %>%
  mutate(Stem_id = as.integer(Stem_id))
cat(sprintf('Traits for S23F8820477:\n'))
print(colnames(S23F8820477)[grep('Adj_',colnames(S23F8820477))])
# Show common traits
# List of data frames
df_list <- list(S22F812A2, S23F812A1, S23F832392, S23F832393, S23F832394, S23F832395, S23F8820476, S23F8820477)  # Add your data frames here
df2 <- list(S23F8820476, S23F8820477)

# Get the column names of the first data frame
common_cols <- colnames(df_list[[1]])

# Iterate through the remaining data frames
for (i in 2:length(df_list)) {
  common_cols <- intersect(common_cols, colnames(df_list[[i]]))
}

# 'common_cols' now contains the column names that are the same across all data frames
print(common_cols)
# Get the column names of the first data frame
common_cols <- colnames(df2[[1]])

# Iterate through the remaining data frames
for (i in 2:length(df_list)) {
  common_cols <- intersect(common_cols, colnames(df_list[[i]]))
}

# 'common_cols' now contains the column names that are the same across all data frames
print(common_cols)
```
## Plots of data 476 and 477
Traits that seems most promising: Hjd_17, Htv_17, Sprant_17, Hjd_07
```{r}
# Height per family
summary_data_476 <- S23F8820476 %>%  
  mutate(Family_id = as.factor(Family_id)) %>%
  group_by(Family_id) %>%
  summarise(
    Avg_Hjd_17 = mean(Hjd_17),
    SD_Hjd_17 = sd(Hjd_17),
    Avg_Hjd_7 = mean(Hjd_07),
    SD_Hjd_7 = sd(Hjd_07), 
  ) %>%
  filter(!is.na(Avg_Hjd_17))
# Choose a ColorBrewer palette (e.g., "Set2")
color_palette <- brewer.pal(nrow(summary_data_476), "Set2")  # Adjust the number as needed

# Create the bar plot
p11 <- ggplot(summary_data_476, aes(x = Family_id)) +
  geom_bar(aes(y = Avg_Hjd_17), stat = "identity", fill = color_palette[1]) +
  scale_fill_manual(values = color_palette) +
  geom_errorbar(aes(ymin = Avg_Hjd_17 - SD_Hjd_17, ymax = Avg_Hjd_17 + SD_Hjd_17), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Family_id", y = "Average Hjd_17", title = "Average and Standard Deviation of Hjd_17 by Family_id for trial F476") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); p11  # Rotate x-axis labels for better readability
p12 <- ggplot(summary_data_476, aes(x = Family_id)) +
  geom_bar(aes(y = Avg_Hjd_7), stat = "identity", fill = color_palette[1]) +
  scale_fill_manual(values = color_palette) +
  geom_errorbar(aes(ymin = Avg_Hjd_7 - SD_Hjd_7, ymax = Avg_Hjd_7 + SD_Hjd_7), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Family_id", y = "Average Hjd_7", title = "Average and Standard Deviation of Hjd_7 by Family_id for trial F476") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); p12  # Rotate x-axis labels for better readability
# Create a histogram of 'Hjd_17'
h1 <- ggplot(S23F8820476, aes(x = Hjd_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Hjd_17", y = "Frequency", title = "Histogram of Hjd_17 for trial F476") +
  theme_minimal(); h1
hht <- ggplot(S23F8820476, aes(x = Htv_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Htv_17", y = "Frequency", title = "Histogram of Htv_17 for trial F476") +
  theme_minimal(); hht
h2 <- ggplot(S23F8820476, aes(x = Hjd_07)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Hjd_7", y = "Frequency", title = "Histogram of Hjd_7 for trial F476") +
  theme_minimal(); h2
h3 <- ggplot(S23F8820476, aes(x = Sgblus_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Sgblus_17", y = "Frequency", title = "Histogram of Sgblus_17 for trial F476") +
  theme_minimal(); h3
h4 <- ggplot(S23F8820476, aes(x = Sprant_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Sprant_17", y = "Frequency", title = "Histogram of Sprant_17 for trial F476") +
  theme_minimal(); h4
cat(sprintf('Checking non-zero values in Sgblus_17 for trial S23F8820476:\n'))
# Count values in 'Sgblus' column that are greater than 0
count_positive_values <- sum(S23F8820476$Sgblus_17 > 0,na.rm = TRUE)
# Print the count
print(count_positive_values)
cat(sprintf('Checking non-zero values in Sprant_17 for trial S23F8820476:\n'))
# Count values in 'Sgblus' column that are greater than 0
count_positive_values <- sum(S23F8820476$Sprant_17 > 0, TRUE)
# Print the count
print(count_positive_values)

summary_data_477 <- S23F8820477 %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  group_by(Family_id) %>%
  summarise(
    Avg_Hjd_17 = mean(Hjd_17),
    SD_Hjd_17 = sd(Hjd_17),
     Avg_Hjd_7 = mean(Hjd_07),
    SD_Hjd_7 = sd(Hjd_07),   
  ) %>%
  filter(!is.na(Avg_Hjd_17))
# Create the bar plot
p1 <- ggplot(summary_data_477, aes(x = Family_id)) +
  geom_bar(aes(y = Avg_Hjd_17), stat = "identity", fill = color_palette[2]) +
  scale_fill_manual(values = color_palette) +
  geom_errorbar(aes(ymin = Avg_Hjd_17 - SD_Hjd_17, ymax = Avg_Hjd_17 + SD_Hjd_17), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Family_id", y = "Average Hjd_17", title = "Average and Standard Deviation of Hjd_17 by Family_id for trial F477") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); p1  # Rotate x-axis labels for better readability
p2 <- ggplot(summary_data_477, aes(x = Family_id)) +
  geom_bar(aes(y = Avg_Hjd_7), stat = "identity", fill = color_palette[2]) +
  scale_fill_manual(values = color_palette) +
  geom_errorbar(aes(ymin = Avg_Hjd_7 - SD_Hjd_7, ymax = Avg_Hjd_7 + SD_Hjd_7), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Family_id", y = "Average Hjd_7", title = "Average and Standard Deviation of Hjd_7 by Family_id for trial F477") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)); p2  # Rotate x-axis labels for better readability
# Create a histogram of 'Hjd_17'
h11 <- ggplot(S23F8820477, aes(x = Hjd_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Hjd_17", y = "Frequency", title = "Histogram of Hjd_17 for trial F477") +
  theme_minimal(); h11
h111 <- ggplot(S23F8820477, aes(x = Htv_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Htv_17", y = "Frequency", title = "Histogram of Htv_17 for trial F477") +
  theme_minimal(); h111
h12 <- ggplot(S23F8820477, aes(x = Hjd_07)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Hjd_7", y = "Frequency", title = "Histogram of Hjd_7 for trial F477") +
  theme_minimal(); h12
h13 <- ggplot(S23F8820477, aes(x = Sgblus_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Sgblus_17", y = "Frequency", title = "Histogram of Sgblus_17 for trial F477") +
  theme_minimal(); h13
h14 <- ggplot(S23F8820477, aes(x = Sprant_17)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = "Sprant_17", y = "Frequency", title = "Histogram of Sprant_17 for trial F477") +
  theme_minimal(); h14
cat(sprintf('Checking non-zero values in Sgblus for trial S23F8820477:\n'))
# Count values in 'Sgblus' column that are greater than 0
count_positive_values <- sum(S23F8820477$Sgblus_17 > 0, na.rm = TRUE)
# Print the count
print(count_positive_values)
cat(sprintf('Checking non-zero values in Sprant_17 for trial S23F8820476:\n'))
# Count values in 'Sprant' column that are greater than 0
count_positive_values <- sum(S23F8820477$Sprant_17 > 0, TRUE)
# Print the count
print(count_positive_values)
```

## Get pedigree based relationship matrix
 There were 8 in-run aggregates extracted from DATAPLAN
 A total of 0 new aggregates have been generated for inclusion in the analysis because there are families where
 1) the male is a group and the female has a non-zero selfing percentage
 2) the male is an aggregate containing the female and the female has a zero selfing percentage
 3) the male is an aggregate not containing the female and female has a non-zero selfing percentage
Number of lines read in            	 22756
Number of genotypes in list        	 22756
Number of genotypes in pedigree    	 22949
Number of genetic groups           	    19
Number of base genotypes           	  3202
Number of families                 	   249
Number of aggregates               	     8
Maximum size of aggregates         	    51
Number of mixes                    	     0
Maximum size of mixes              	     0
Number of parents                  	   193
lastgen 1
Generation Count Average F
----------------------------------
     0     3202   0.0000
     1     19747  0.0002
----------------------------------
           22949

Because there are aggregates and/or mixes in the pedigree a giv file has been generated.
The pedigree file contains '0' for any dad that is an aggregate
You will need to specify this giv file in any ASReml analysis that uses this pedigree


To fit a GIV matrix in ASReml use the following syntax

	ped.giv !SKIP 1 - specified on the line immediately prior to the data file line
	giv(Genotype_id,1) - in the model statement for univariate analysis
	Genotype_id 0 GIV - in the G structure definition line for multivariate analysis
```{r}
###### get all genotype ids and save to file
genotype_list <- data.frame(Genotype_id = c(S22F812A2$Genotype_id,S23F812A1$Genotype_id,S23F832392$Genotype_id,S23F832393$Genotype_id,S23F832394$Genotype_id,S23F832395$Genotype_id,S23F8820476$Genotype_id,S23F8820477$Genotype_id))
write.table(x = genotype_list,file = paste(path,"genotype_id.txt",sep=""),quote = FALSE,row.names = F)
pedigree <- read.table(file = paste(path,"pedigree.txt",sep=""),sep=",",header=T)
ainv <- read.table(file = paste(path,"ped.giv",sep=""),sep=",",header=T)


genotype_list2 <- data.frame(Genotype_id = c(S23F8820476$Genotype_id,S23F8820477$Genotype_id))
write.table(x = genotype_list2,file = paste(path,"genotype_id_subs.txt",sep=""),quote = FALSE,row.names = F)
```


## Only two trials 476 and 477
There were 4 in-run aggregates extracted from DATAPLAN
 A total of 0 new aggregates have been generated for inclusion in the analysis because there are families where
 1) the male is a group and the female has a non-zero selfing percentage
 2) the male is an aggregate containing the female and the female has a zero selfing percentage
 3) the male is an aggregate not containing the female and female has a non-zero selfing percentage
Number of lines read in            	  5992
Number of genotypes in list        	  5992
Number of genotypes in pedigree    	  6151
Number of genetic groups           	    13
Number of base genotypes           	   979
Number of families                 	   146
Number of aggregates               	     4
Maximum size of aggregates         	    51
Number of mixes                    	     0
Maximum size of mixes              	     0
Number of parents                  	   159
lastgen 1
Generation Count Average F
----------------------------------
     0     979    0.0000
     1     5172   0.0005
----------------------------------
           6151 

Because there are aggregates and/or mixes in the pedigree a giv file has been generated.
The pedigree file contains '0' for any dad that is an aggregate
You will need to specify this giv file in any ASReml analysis that uses this pedigree


To fit a GIV matrix in ASReml use the following syntax

	ped.giv !SKIP 1 - specified on the line immediately prior to the data file line
	giv(Genotype_id,1) - in the model statement for univariate analysis
	Genotype_id 0 GIV - in the G structure definition line for multivariate analysis



Because you have incorporated 13 genetic groups, ped.giv has '!GROUPSDF 13' on 2nd line

```{r}
pedigree2 <- read.table(file = paste(path,"pedigree (2).txt",sep=""),sep=",",header=T)
cat(sprintf('Number of individuals in pedigree: %d\n',nrow(pedigree2)))
# ,!GROUPSDF 13
ainv2 <- read.table(file = paste(path,"ped (2).giv",sep=""),sep=",",header=T)
##### merging dataset
data_476 <- S23F8820476 %>%
  select(c(Stem_id,Prow,Ppos,Ruta, Rad, Planta, Genotype_id,Family_id, Mum_id, Mum_type, Dad_id, Dad_type, Hjd_17, Htv_17, Sprant_17, Vit_07, Hjd_07)) %>%
  mutate(Trial = "S23F8820476")
data_477 <- S23F8820477 %>%
  select(c(Stem_id,Prow,Ppos, Ruta, Rad, Planta, Genotype_id,Family_id, Mum_id, Mum_type, Dad_id, Dad_type, Hjd_17, Htv_17, Sprant_17, Vit_07, Hjd_07)) %>%
  mutate(Trial = "S23F8820477")
data_476_477 <- rbind(data_476,data_477)
rm(data_476, data_477)
# Filter and save the Genotype_id of individuals to be removed
pedigree3 <- pedigree2 %>% # remove checklot trees
  filter(!(Dad_id == 0) | grepl("Picea_abies", Genotype_id)) %>% #  & dam == 0
  mutate(Genotype_id = as.factor(Genotype_id))

# Filter 'data_476_477' to keep only individuals that match the retained Genotype_id values
data_476_477_wo_check <- data_476_477 %>%
  filter(Genotype_id %in% pedigree3$Genotype_id)
```




## Check proportion of genotyped individuals within trials

```{r}
count_overlap <- function(df1,df2){
  
  #df1 <- data.frame(Stem_id = S22F812A2$Stem_id)
  #df2 <- data.frame(ID = id$ID)
  matching_indices <- match(df1, df2, nomatch = 0)

# Count the non-zero matches
  intersection_count <- sum(matching_indices > 0)

# 'intersection_count' now contains the count of matching elements
  print(intersection_count)
  #intersection_count <- df1 %>%
  #  inner_join(df2, by = c("Stem_id" = "ID")) %>%
  #  nrow()
  #intersection_count <-  length(intersect(df1, df2))
  # 'intersection_count' now contains the count of matching elements
  print(intersection_count)
  return(intersection_count)
}

id <-read.table(paste(path,"ID_genotyped.txt",sep=""),sep = ",",header=T) %>%
  select(ID) %>%
  as.data.frame() %>%
  mutate(ID = as.integer(ID))
  

count_overlap(S23F812A1$Genotype_id,id$ID)
count_overlap(S22F812A2$Genotype_id,id$ID)
count_overlap(S23F832392$Genotype_id,id$ID)
count_overlap(S23F832393$Genotype_id,id$ID)
count_overlap(S23F832394$Genotype_id,id$ID)
count_overlap(S23F832395$Genotype_id,id$ID)
count_overlap(S23F8820476$Genotype_id,id$ID)
count_overlap(S23F8820477$Genotype_id
              ,id$ID)
```



## Read relationship matrix

```{r}
G <- read.csv(file = "Gmat_rrBLUP.csv",header = T,check.names = F,row.names = 1)
print(dim(G))
G <- as.matrix(G)
G <- round(G,3) # improve numerical stability

#snp.data <- read.csv("markers_imputed.csv", row.names = 1)

# pheatmap(G, main = "G matrix",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0)

```
## Checking G - A matrices

```{r}
# snp data
snp.sol <- read.table(file = "Save/genotype_data_1218_JWAS.csv",sep=",",header = TRUE) # ,row.names = TRUE
#Computing the numerator relationship matrix 10% of double-reduction
Amat <- Amatrix(ped.sol, ploidy=2, w = 0.1)
Gmat <- Gmatrix(snp.sol, ploidy=2,
                maf=0.05, method="VanRaden")
Gmat <- round(Gmat,3) #see appendix

#Computing H matrix (Martini)
Hmat_Martini <- Hmatrix(A=Amat, G=G, method="Martini", 
                        missingValue=-9, maf=0.05)

#Computing H matrix (Munoz)
Hmat_Munoz <- Hmatrix(A=Amat, G=Gmat, markers = snp.sol, method="Munoz", 
                      missingValue=-9, maf=0.05)
```

## Calculate H matrix
H matrix is the relationship matrix using combined information from the pedigree and genomic relationship matrices. First, you need to compute the matrices separated and then use them as input to build the combined H matrix. 
```{r}
pedigree3 <- pedigree3 %>%
  rename(sire = Dad_id, dam = Mum_id)
pedigree4 <- pedigree3 %>%
  filter(sire != 0 & dam != 0) %>%
  mutate(
    Sire = ifelse(grepl("\\[", sire), 0, sire),
    Dam = ifelse(grepl("\\[", dam), 0, dam)
  ) %>%
  rename(ID = Genotype_id)
 # select(c(Genotype_id,Mum_id,Dad_id)) %>%
 # rename(dam = Mum_id, sire = Dad_id) # ID = Individual 

Amat <- as.matrix(nadiv::makeA(pedigree4[,c("ID","Dam","Sire")])) # Create A mat
print(dim(Amat))

ro <- intersect(rownames(Amat), rownames(G))
co <- intersect(colnames(Amat), colnames(G))
A22 <- Amat[ro,co]
G22 <- G[ro,co]
pheatmap(A22, main = "A matrix Picea abies 1218 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0)
pdf(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/Figures/Amatrix_1218.pdf")
pheatmap(A22, main = "A matrix Picea abies 1218 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0)
dev.off()
png(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/Figures/Amatrix_1218.png")
pheatmap(A22, main = "A matrix Picea abies 1218 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0)
dev.off()
pheatmap1 <- pheatmap(A22, main = "A matrix Picea abies 1218 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0)
row_order <- pheatmap1$tree_row$order
col_order <- pheatmap1$tree_col$order
Gmat2 <- G22[row_order, col_order]
pheatmap(Gmat2, main = "G matrix Picea abies 1218 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0, cluster_rows = FALSE, cluster_cols = FALSE)
###################### save
pdf(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/Figures/Gmatrix_1218.pdf")
pheatmap(Gmat2, main = "G matrix taeda 926 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0, cluster_rows = FALSE, cluster_cols = FALSE)
dev.off()
png(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/Figures/Gmatrix_1218.png")
pheatmap(Gmat2, main = "G matrix taeda 926 genotypes",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0, cluster_rows = FALSE, cluster_cols = FALSE)
dev.off()
################ write data to file
# fix order and check overlap
# save to file
write.table(x = A22,file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/TaedaPineData/results_A_926/Amatrix_spruce_1218_ordered.txt",sep = ",",quote = FALSE,col.names = FALSE,row.names = FALSE)
write.table(x = ordered_matrix,file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/TaedaPineData/results_A_926/Amatrix_spruce_1218_ordered_names.txt",sep = ",",quote = FALSE,col.names = FALSE,row.names = TRUE)


#write.table(x = A22,file = "Save/Amat_1218.txt",sep="\t",quote = FALSE,row.names = TRUE)
# make.positive.definite computes the nearest positive definite of a real symmetric matrix, using the algorithm of NJ Higham (1988)
#cat(sprintf("Check G matrix for positive definiteness.....\n"))
#print(is.positive.definite(G22))
#print(rank.condition(G22)) # numerically ill-conditioned?
#cat(sprintf("Make G22 positive definite.....\n"))
#G23 <- make.positive.definite(G22) # make positive definite
#print(is.positive.definite(G23))
#print(rank.condition(G23)) 
#print(G23[1:10,1:10]-G22[1:10,1:10])
#write.table(x = G23,file = "Save/Gmat_1218.txt",sep="\t",quote = FALSE,row.names = TRUE)
#pheatmap(A22, main = "A22 matrix",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0,filename = "Figures/heatmap_A22.png")
#pheatmap(G22, main = "G matrix",show_rownames = FALSE, show_colnames = FALSE,treeheight_row = 0, treeheight_col = 0,filename = "Figures/heatmap_G.png")
#Computing H matrix (Martini)
#G23_0 <- G23

#index <- which(G23_0[,]<0) 
#G23_0[index] <- 0
#print(is.positive.definite(G23_0))
Hmat <- Hmatrix(A=Amat, G=G, method="Martini", tau = 1, omega = 1)
print(dim(Hmat))
#Computing H matrix (Munoz)
#Hmat_Munoz <- Hmatrix(A=Amat, G=G, 
#                      ploidy=2, method="Munoz",markers = snp.data) # markers = snp.sol, 


# data_476_477_wo_check
added_ids <- setdiff(colnames(Hmat),data_476_477_wo_check$Genotype_id)

# Assuming 'added_ids' contains the row and column names to be removed
Hmat2 <- Hmat[setdiff(rownames(Hmat), added_ids), setdiff(colnames(Hmat), added_ids)]
Amat2 <- Amat[setdiff(rownames(Amat), added_ids), setdiff(colnames(Amat), added_ids)]
## Save these for use in OSQP and Opsel
#write.table(x = Hmat2,file = "Save/Hmat_5525.txt",sep="\t",quote = FALSE,row.names = TRUE)
#write.table(x = Amat2,file = "Save/Amat_5525.txt",sep="\t",quote = FALSE,row.names = TRUE)
############# add Lev_7 and Lev_17
data_476_477_wo_check <- data_476_477_wo_check %>%
  mutate(Lev_17 = ifelse(!is.na(Hjd_17) & Hjd_17 > 0, 1, 0)) %>%
  mutate(Lev_7 = ifelse(!is.na(Hjd_07) & Hjd_07 > 0, 1, 0))

#############
order_index <- match(colnames(Hmat2), data_476_477_wo_check$Genotype_id)
# Use the index to reorder the rows of the data frame
data_476_477_wo_check1 <- data_476_477_wo_check[order_index, ]
print(dim(data_476_477_wo_check1))
######### save data for analysis in Julia, JWAS package
data_tmp <- data_476_477_wo_check1 %>%
  select(Genotype_id,Rad,Ruta,Planta,Trial,Mum_id,Dad_id,Family_id,Hjd_17,
         Hjd_07,Sprant_17,Htv_17,Vit_07,Lev_7,Lev_17) %>%
  mutate(Genotype_id = paste("I",Genotype_id,sep="")) %>%
  mutate(Mum_id = paste("I",Mum_id,sep="")) %>%
  mutate(Dad_id = paste("I",Dad_id,sep = "")) %>%
  mutate(Family_id = paste("I",Family_id,sep = "")) %>%
  mutate(Trial_Family = interaction(Trial, Family_id, sep = "_")) %>%
  mutate(Trial_Ruta = interaction(Trial, Ruta, sep = "_"))
write.table(x = data_tmp,file = "Save/phenotypes_5525_spruce_Horder_v2.txt",sep=",",row.names = FALSE,quote = FALSE)
Hmat_tmp <- Hmat2
colnames(Hmat_tmp) <- rownames(Hmat_tmp) <- paste("I",colnames(Hmat2),sep = "")
# make.positive.definite computes the nearest positive definite of a real symmetric matrix, using the algorithm of NJ Higham (1988)
Hmat_tmp <- make.positive.definite(Hmat_tmp) # make positive definite
write.table(x = round(Hmat_tmp,digits = 3),file = "Save/Hmat_5525_spruce.txt",row.names = TRUE,quote = FALSE,sep=",")
# Round the matrix to three decimal places
Hmat_tmp_rounded <- round(Hmat_tmp, digits = 3)

# Convert the matrix to a data frame (optional, but recommended for better formatting)
Hmat_tmp_df <- as.data.frame(Hmat_tmp_rounded)

# Write the data frame to a CSV file
write.csv(Hmat_tmp_df, file = "Save/Hmat_5525_spruce.csv", row.names = TRUE, quote = FALSE)
write.table(Hmat_tmp_df, file = "Save/Hmat_5525_spruce_v2.txt", sep = ",", row.names = TRUE, col.names = FALSE, quote = FALSE)
write(x = Hmat_tmp_rounded,file = "Save/Hmat_5525_spruce_v3.txt",ncolumns = ncol(Hmat_tmp_rounded),sep = ",")
require(MASS)
write.matrix(x = Hmat_tmp,file = "Save/Hmat_5525_spruce_v2.txt",sep = ",",)
rm(Hmat_tmp)
print(dim(Hmat2))
print(head(data_476_477_wo_check1$Genotype_id,n = 30))
print(head(rownames(Hmat2),n = 30))
print(head(colnames(Hmat2),n = 30))
########### write pedigree to JWAS package
write.table(x = data.frame(ID = data_tmp$Genotype_id, sire = data_tmp$Dad_id, dam = data_tmp$Mum_id),file = "Save/pedigree_H_5525.txt",quote = FALSE,sep=",",row.names = FALSE)



order_index <- match(colnames(Amat2), data_476_477_wo_check$Genotype_id)
# Use the index to reorder the rows of the data frame
data_476_477_wo_check2 <- data_476_477_wo_check[order_index, ]
print(dim(data_476_477_wo_check2))
print(dim(Amat2))
print(head(data_476_477_wo_check2$Genotype_id,n = 30))
print(head(rownames(Amat2),n = 30))
print(head(colnames(Amat2),n = 30))
######### save data for analysis in Julia, JWAS package
data_tmp <- data_476_477_wo_check2 %>%
  select(Genotype_id,Rad,Ruta,Planta,Trial,Mum_id,Dad_id,Family_id,Hjd_17,Hjd_07,Sprant_17,Htv_17,Vit_07,Lev_7,Lev_17)  %>%
  mutate(Genotype_id = paste("I",Genotype_id,sep="")) %>%
  mutate(Mum_id = paste("I",Mum_id,sep="")) %>%
  mutate(Dad_id = paste("I",Dad_id,sep = "")) %>%
  mutate(Family_id = as.factor(paste("I",Family_id,sep = ""))) %>%
  mutate(Trial_Family = interaction(data_tmp$Trial, data_tmp$Family_id, sep = "_")) %>%
  mutate(Trial_Ruta = interaction(Trial, Ruta, sep = "_"))

write.table(x = data_tmp,file = "Save/phenotypes_5525_spruce_Aorder.txt",sep=",",row.names = FALSE,quote = FALSE)
Amat_tmp <- Amat2
colnames(Amat_tmp) <- rownames(Amat_tmp) <- paste("I",colnames(Amat2),sep = "")
write.table(x = Amat_tmp,file = "Save/Amat_5525_spruce.txt",row.names = TRUE,quote = FALSE,sep=",")
rm(Amat_tmp)
write.table(x = data_tmp,file = "Save/phenotypes_5525_spruce_Aorder.txt",sep=",",row.names = FALSE,quote = FALSE)
#write.table(x = Amat2,file = "Save/Amat_5525_spruce.txt",row.names = TRUE,quote = FALSE,sep=",")

# Assuming your data frame is named data_476_477_wo_check
data_476_477_wo_check_na <- data_476_477_wo_check %>%
  bind_rows(data.frame(Genotype_id = added_ids), .)

##### H
# Create an index based on the order of the external vector
order_index <- match(colnames(Hmat), data_476_477_wo_check_na$Genotype_id)
# Use the index to reorder the rows of the data frame
data_476_477_wo_check_naH <- data_476_477_wo_check_na[order_index, ]
print(dim(data_476_477_wo_check_naH))
print(dim(Hmat))
print(head(data_476_477_wo_check_naH$Genotype_id,n = 30))
print(head(rownames(Hmat),n = 30))
print(head(colnames(Hmat),n = 30))
# Create an index based on the order of the external vector
order_index <- match(colnames(Amat), data_476_477_wo_check_na$Genotype_id)
#### A
# Use the index to reorder the rows of the data frame
data_476_477_wo_check_naA <- data_476_477_wo_check_na[order_index, ]
print(dim(data_476_477_wo_check_naA))
print(dim(Amat))
print(head(data_476_477_wo_check_naA$Genotype_id,n = 30))
print(head(rownames(Amat),n = 30))
print(head(colnames(Amat),n = 30))

####### order ok and missing data added to founders


```

## Save to Opsel format

```{r}
# Save data for Opsel analysis
# column names: Indiv.,Female,Male,EBV,Max
# First read in results from JWAS analysis to get EBV
EBV_Hjd_17 <- read.table(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/results_JWAS_5525_A/EBV_Hjd_17.txt",sep=",",header = TRUE)
EBV_Htv_17 <- read.table(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/results_JWAS_5525_A/EBV_Htv_17.txt",sep=",",header = TRUE)
EBV_Sprant_17 <- read.table(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/results_JWAS_5525_A/EBV_Sprant_17.txt",sep=",",header = TRUE)
EBV_Vit_7 <- read.table(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/results_JWAS_5525_A/EBV_Vit_07.txt",sep=",",header = TRUE)
EBV_Hjd_7 <- read.table(file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/results_JWAS_5525_A/EBV_Hjd_07.txt",sep=",",header = TRUE)
# Index 0.6 growth, 0.2 Quality, 0.2 Vitality
EBV_Index <- EBV_Hjd_17 %>%
  mutate(EBV = (0.25*scale(EBV_Hjd_17$EBV)[,1] + 0.25*scale(EBV_Htv_17$EBV)[,1] + 0.1*scale(EBV_Hjd_7$EBV)[,1] + 0.2*scale(EBV_Sprant_17$EBV)[,1] + 0.2*scale(EBV_Vit_7$EBV)[,1])) %>%
  select(-c(PEV))
# Add parents to pedigree
# find unique numbers 
parents <- unique(c(data_tmp$Dad_id,data_tmp$Mum_id))
f0 <- data.frame(Indiv. = parents, Female = 0,Male = 0,EBV = rep(0,length(parents)),Max = rep(0,length(parents)))
f1 <- data.frame(Indiv. = data_tmp$Genotype_id,Female = data_tmp$Mum_id,Male = data_tmp$Dad_id,EBV = EBV_Index$EBV,Max = rep(5,5525))
df <- rbind(f0,f1)
# Indiv.,Female,Male,EBV,Max
write.table(x = df,file = "Save/Input_opsel_5525_selindex.txt",quote= FALSE, sep=",",row.names = FALSE)
```


## Calculate G inverse

```{r}
dim<-dim(X)
allfreq<-colSums(X)/(dim[1]*2)
#Only if not already pruned for MAF
#freqdel1<-which(allfreq<0.01)
#freqdel2<-which(allfreq>0.99)
#X<-X[,-c(freqdel1,freqdel2)]
#allfreq<-allfreq[-c(freqdel1,freqdel2)]
m<-length(allfreq)

#Calculate G matrix and its inverse
M<-X-1
Z<-sweep(M,2,2*(allfreq-0.5),"-")
G<-(Z%*%t(Z))/(2*sum(allfreq*(1-allfreq)))
library(corpcor)
library(INLA)
G<-make.positive.definite(G)
Ginv<-as(as(as(solve(G), "dMatrix"), "generalMatrix"), "TsparseMatrix")
Ginv[ abs(Ginv) < sqrt(.Machine$double.eps) * max(Ginv) ] = 0

#Free memory
rm(dat)
rm(M)
rm(Z)
rm(G)

```

## Stan via brms
We use bmrs which runs Bayesian Multilevel Models using Stan. NB we need to carefully select an adequate prior for the analysis to improve inference. 
```{r}

#### Define predictor set
# Traits: Hjd_17, Htv_17, Sprant_17, Hjd_07, Vit_07
# 
bf_Hjd17 <- bf(Hjd_17 ~ 1 + Trial + (1 | a | gr(Genotype_id, cov = Hmat)) + (1 | b | Trial:Family_id) + (1 | c | Trial:Ruta)) # + (1 | d | Ruta*Rad) + (1 | d | Ruta*Planta) + (1 | e | Trial*Family_id)
bf_vit7 <- bf(Vit_07 ~ 1 + Trial + (1 | a | gr(Genotype_id, cov = Hmat)) + (1 | b | Trial:Family_id)  + (1 | c | Trial:Ruta))
bf_hjd7 <- bf(Hjd_07 ~ 1 + Trial + (1 | a | gr(Genotype_id, cov = Hmat)) + (1 | b | Trial:Family_id)  + (1 | c | Trial:Ruta))
bf_Sprant17 <- bf(Sprant_17 ~ 1 + Trial + (1 | a | gr(Genotype_id, cov = Hmat)) + (1 | b | Trial:Family_id)  + (1 | c | Trial:Ruta))
bf_Htv_17 <- bf(Htv_17 ~ 1 + Trial + (1 | a | gr(Genotype_id, cov = Hmat)) + (1 | b | Trial:Family_id)  + (1 | c | Trial:Ruta))
brms_m1.1 <- brm(
  bf_Hjd17, # bf_hjd7 + bf_Htv_17 + #  + bf_vit7 + bf_Sprant17 + set_rescor(TRUE)
  data = data_476_477_wo_check1,
  data2 = list(Hmat = Hmat2),
  family = gaussian(link = "identity"), # c("gaussian", "gaussian", "poisson"), #gaussian(), gaussian(link = "identity"), categorical(link = "logit", refcat = NULL)
  chains = 4,  # 6
  cores = 4,  # 6
  iter = 700, # 1000
  warmup = 200, # 500
  thin = 1
)

save(brms_m1.1, file = "C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Optimum contribution selection/NorwaySpruceData/Save/brms_m1_NorwaySpruce_hjd17_Hmat.rda")
brms_m1.1 <-load("Save/brms_m1_NorwaySpruce_hjd17_Hmat.rda")
plot(brms_m1.1)
summary(brms_m1.1)
mcmc_plot(brms_m1.1, type = "acf")
###################### heritability ###############################
v_Individual <- (VarCorr(brms_m1.1, summary = FALSE)$Genotype_id$sd)^2
v_Family <- (VarCorr(brms_m1.1, summary = FALSE)$'Trial:Family_id'$sd)^2 # NB Check
v_Ruta <- (VarCorr(brms_m1.1, summary = FALSE)$'Trial:Ruta'$sd)^2
v_r <- (VarCorr(brms_m1.1, summary = FALSE)$residual$sd)^2

h.HT17 <- as.mcmc(v_Individual[, 1] / (v_Individual[, 1] + v_Family[,1] + v_Ruta[,1] + v_r[, 1])) # NB add family variance in denominator
#h.HT2.2 <- as.mcmc(v_Individual[, 2] / (v_Individual[, 2] + v_r[, 2])) # NB check index
#h.HT3.2 <- as.mcmc(v_Individual[, 3] / (v_Individual[, 3] + v_r[, 3])) # NB check index

summary(h.HT17)
#summary(h.HT2.2)
#summary(h.HT3.2)
plot(h.HT17)
#plot(h.HT2.2)
#plot(h.HT3.2)
#Var.table <- as_draws_df(brms_m1.1)
#Var.table$h.bwt.1 <- as.mcmc((Var.table$sd_Individual__Intercept)^2 / ((Var.table$sd_Individual__Intercept)^2 + (Var.table$sigma)^2))
#summary(Var.table$h.bwt.1)
#plot(Var.table$h.bwt.1)

############################## Correlations ###########################
#cor_g_12 <- as.mcmc((VarCorr(brms_m1.1, summary = FALSE)$Genotype_id$cor[, 1, 2]))
#cor_g_13 <- as.mcmc((VarCorr(brms_m1.1, summary = FALSE)$Genotype_id$cor[, 1, 3]))
#cor_res <- as.mcmc((VarCorr(brms_m1.1, summary = FALSE)$residual$cor[, 1, 2]))
#cor_family <- as.mcmc((VarCorr(brms_m1.1, summary = FALSE)$Family$cor[, 1, 2]))
#summary(cor_g_12)
#summary(cor_g_13)
#summary(cor_res)
#summary(cor_family)


# Summary of the model
summary(brms_m1.1)

# Plot the model
plot(brms_m1.1)
```

## BGLR
The BGLR (‘Bayesian Generalized Linear Regression’) function fits various types of parametric and semi-parametric Bayesian regressions to continuos (censored or not), binary and ordinal outcomes.
```{r}
# Both A and H matrices
#niter = 15000
#burnin = 5000

#  BGLR(y, response_type = "gaussian", a=NULL, b=NULL,ETA = NULL, nIter = 1500,
#       burnIn = 500, thin = 5, saveAt = "", S0 = NULL, 
#       df0 =5, R2 = 0.5, weights = NULL,
#       verbose = TRUE, rmExistingFiles = TRUE, groups=NULL)


#2# Computing the distance matrix and then the krenel.
#X<-scale(X,center=TRUE,scale=TRUE)
#D<-(as.matrix(dist(X,method='euclidean'))^2)/p
#h<-0.5
#K<-exp(-h*D)

##### pedigree NRM for comparison with Opsel
y <- data_476_477_wo_check2$Hjd_17
#Trial <- data_476_477_wo_check$Trial
data_476_477_wo_check2 <- data_476_477_wo_check2 %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  mutate(Trial = as.factor(Trial))
X <- model.matrix(~ Trial, data = data_476_477_wo_check2)
Z1 <- model.matrix(~ Trial:Ruta, data_476_477_wo_check2)
Z2 <-  model.matrix(~ Trial:Family_id, data_476_477_wo_check2)
#3 setting up the linear predictor
ETA<-list( 
  list(X = X,model='FIXED',saveEffects=TRUE),
  list(X = Z1,model="BRR",saveEffects=TRUE),
  list(X = Z2,model="BRR",saveEffects=TRUE),
  list(K=Amat2, model='BRR',saveEffects=TRUE) # BR
)
# Fitting the model
### Runs the Gibbs sampler
#fm1<-BLR(y=y, XL=X,XF = Z1,XR = Z2,GF=list(ID=1:nrow(Amat2),A=Amat2),
#                           nIter=5500,burnIn=500,thin=1, saveAt = #"Save/saverun_BLR_A_Hjd17")
fm<-BGLR(y=y,ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_A_Hjd17") # ,groups = data_476_477_wo_check$Trial
save(fm,file='Save/fm_A_hjd17.rda')
# Indiv.,Female,Male,EBV,Max format needed for Opsel
res.Hjd17 <- data.frame(Id = data_476_477_wo_check2$Genotype_id,ID2 = rownames(Amat2), Dad = data_476_477_wo_check2$Dad_id, Mum = data_476_477_wo_check2$Mum_id, EBV = fm$ETA[[4]]$u, StdErr = fm$ETA[[4]]$SD.u, Max = 10)
write.table(x = res.Hjd17,file = "Save/res_BGLR_Amat_5525_Hjd17.txt",quote = FALSE,row.names = TRUE)
summary(fm)
##Vulcano plot (posterior SD vs estimated effects)
png(filename = "Figures/vulcano_plot_Amat_hjd17.png")
plot(fm$ETA[[4]]$u~fm$ETA[[4]]$SD.u,col=2,
 main='Vulcano Plot (continuous outcome)',
 xlab='Estimated Effect',ylab='Est. Posterior SD')
dev.off()
png(filename = "Figures/EBV_hjd17_Amat.png")
plot(fm$ETA[[4]]$u~y,col=3,
 main='EBV vs Hjd_17 (continuous outcome)',
 xlab='Phenotype',ylab='Estimated Effect')
dev.off()
# Residual variance
print(fm$fit)
varE<-scan(file = 'varE.dat')
png(filename = "Figures/resid_Amat_Hjd17.png")
plot(varE,type='o',col=2,cex=.5,ylab=expression(var[e]));
abline(h=fm$varE,col=4,lwd=2);
abline(v=fm$burnIn/fm$thin,col=4)
dev.off()
##### h2
h2 <- fm$ETA[[4]]$varU/(fm$ETA[[4]]$varU+var(fm$ETA[[2]]$b)+var(fm$ETA[[3]]$b)+fm$varE)
cat(sprintf('h2 = %f\n',h2))
# Total prediction
yHat<-fm$yHat
tmp<-range(c(y,yHat),na.rm = TRUE)
png(filename = "Figures/Predicted_vs_y_Hmat_hjd17.png")
plot(yHat~y,xlab='Observed',ylab='Predicted',col=2,
xlim=tmp,ylim=tmp); abline(a=0,b=1,col=4,lwd=2)
dev.off()

#2# Ordinal outcome
#yOrd<-ifelse(y<quantile(y,1/4),1,ifelse(y<quantile(y,3/4),2,3))
#yOrdNA<-yOrd ; yOrdNA[tst]<-NA
yOrd <- data_476_477_wo_check2$Sprant_17
##### check distribution of Sprant_17
data_476_477_wo_check2 %>% 
  count(Sprant_17)
fmOrd<-BGLR(y=yOrd,response_type='ordinal',ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_A_Sprant17")
save(fmOrd,file='Save/fm_A_sprant17.rda')
head(fmOrd$probs)
summary(fmOrd)
##Vulcano plot (posterior SD vs estimated effects)
png(filename = "Figures/vulcano_plot_Amat_Sprant17.png")
plot(fmOrd$ETA[[4]]$u~fmOrd$ETA[[4]]$SD.u,col=2,
 main='Vulcano Plot (continuous outcome)',
 xlab='Estimated Effect',ylab='Est. Posterior SD')
dev.off()
png(filename = "Figures/EBV_sprant17_Amat.png")
plot(fmOrd$ETA[[4]]$u~y,col=3,
 main='EBV vs Sprant_17 (ordinal outcome)',
 xlab='Phenotype',ylab='Estimated Effect')
dev.off()
# Residual variance
print(fmOrd$fit)
varE<-scan(file = 'varE.dat')
png(filename = "Figures/resid_Amat_Sprant17.png")
plot(varE,type='o',col=2,cex=.5,ylab=expression(var[e]));
abline(h=fmOrd$varE,col=4,lwd=2);
abline(v=fmOrd$burnIn/fm$thin,col=4)
dev.off()
##### h2
h2 <- fmOrd$ETA[[4]]$varU/(fmOrd$ETA[[4]]$varU+var(fmOrd$ETA[[2]]$b)+var(fmOrd$ETA[[3]]$b)+fmOrd$varE)
cat(sprintf('h2 = %f\n',h2))
# Total prediction
yHat<-fmOrd$yHat
tmp<-range(c(yOrd,yHat),na.rm = TRUE)
png(filename = "Figures/Predicted_vs_y_Amat_sprant17.png")
plot(yHat~yOrd,xlab='Observed',ylab='Predicted',col=2,
xlim=tmp,ylim=tmp); abline(a=0,b=1,col=4,lwd=2)
dev.off()
## Lev_17
# step one, create Lev_17
data_476_477_wo_check2 <- data_476_477_wo_check2 %>%
  mutate(Lev_17 = if_else(Hjd_17 > 0, 1, 0)) %>%
  mutate(Lev_17 = ifelse(is.na(Lev_17), 0, Lev_17))
##### check distribution of Lev_17
data_476_477_wo_check2 %>% 
  count(Lev_17)
yBin <- data_476_477_wo_check$Lev_17
fmBin<-BGLR(y=yBin,response_type='ordinal',ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_A_Lev17")
save(fmBin,file='Save/fm_A_lev17.rda')
head(fmBin$probs)

# Create selection index based on 60% production, 20% vitality and 20% quality



####### H matrix model
y <- data_476_477_wo_check1$Hjd_17
#Trial <- data_476_477_wo_check$Trial
data_476_477_wo_check1 <- data_476_477_wo_check1 %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  mutate(Trial = as.factor(Trial))
X <- model.matrix(~ Trial, data = data_476_477_wo_check1)
Z1 <- model.matrix(~ Trial:Ruta, data_476_477_wo_check1)
Z2 <- model.matrix(~ Trial:Family_id, data_476_477_wo_check1)
#3 setting up the linear predictor
ETA<-list( 
  list(X = X,model='FIXED',saveEffects=TRUE),
  list(X = Z1,model='BRR',saveEffects=TRUE),
  list(X = Z2,model='BRR',saveEffects=TRUE),
  list(K=Hmat2, model='BRR',saveEffects=TRUE)
)
#3# Fitting the model
fm22<-BLR(y=y, XL=X,XF = Z1,XR = Z2,GF=list(ID=1:nrow(Hmat2),A=Hmat2),
                           nIter=5500,burnIn=500,thin=1, saveAt = "Save/saverun_BLR_G_Hjd17")
save(fm22,file='Save/fm_BLR_H_hjd17.rda')
fm2<-BGLR(y=y,ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_H_Lev17") # ,groups = data_476_477_wo_check$Trial
save(fm2,file='Save/fm_H_hjd17.rda')
load(file = 'Save/fm_H_hjd17.rda')
#U=readBinMat('ETA_1_b.bin')
summary(fm2)
# Indiv.,Female,Male,EBV,Max format needed for Opsel
res.Hjd17 <- data.frame(Id = data_476_477_wo_check1$Genotype_id,ID2 = rownames(Hmat2), Dad = data_476_477_wo_check1$Dad_id, Mum = data_476_477_wo_check1$Mum_id, EBV = fm2$ETA[[4]]$u, StdErr = fm2$ETA[[4]]$SD.u, Max = 10)
write.table(x = res.Hjd17,file = "Save/res_BGLR_Hmat_5525_Hjd17.txt",quote = FALSE,row.names = TRUE)
##Vulcano plot (posterior SD vs estimated effects)
png(filename = "Figures/vulcano_plot_Hmat_hjd17.png")
plot(fm$ETA[[3]]$u~fm$ETA[[3]]$SD.u,col=2,
 main='Vulcano Plot (continuous outcome)',
 xlab='Estimated Effect',ylab='Est. Posterior SD')
dev.off()
png(filename = "Figures/EBV_hjd17_Hmat.png")
plot(fm$ETA[[3]]$u~y,col=3,
 main='EBV vs Hjd_17 (continuous outcome)',
 xlab='Phenotype',ylab='Estimated Effect')
dev.off()
# Residual variance
print(fm$fit)
varE<-scan(file = 'varE.dat')
png(filename = "Figures/resid_Hmat_Hjd17.png")
plot(varE,type='o',col=2,cex=.5,ylab=expression(var[e]));
abline(h=fm$varE,col=4,lwd=2);
abline(v=fm$burnIn/fm$thin,col=4)
dev.off()
##### h2
h2 <- fm2$ETA[[4]]$varU/(fm2$ETA[[4]]$varU+var(fm2$ETA[[2]]$b)+var(fm2$ETA[[3]]$b)+fm2$varE)
cat(sprintf('h2 = %f\n',h2))
# Total prediction
yHat<-fm2$yHat
tmp<-range(c(y,yHat),na.rm = TRUE)
png(filename = "Figures/Predicted_vs_y_Hmat_hjd17.png")
plot(yHat~y,xlab='Observed',ylab='Predicted',col=2,
xlim=tmp,ylim=tmp); abline(a=0,b=1,col=4,lwd=2)
dev.off()

#2# Ordinal outcome
#yOrd<-ifelse(y<quantile(y,1/4),1,ifelse(y<quantile(y,3/4),2,3))
#yOrdNA<-yOrd ; yOrdNA[tst]<-NA
yOrd <- data_476_477_wo_check1$Sprant_17
fmOrd2<-BGLR(y=yOrd,response_type='ordinal',ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_H_Sprant17") # ,groups = data_476_477_wo_check$Trial
save(fmOrd2,file='Save/fm_H_sprant17.rda')
head(fmOrd2$probs)
##### h2
h2 <- fmOrd2$ETA[[4]]$varU/(fmOrd2$ETA[[4]]$varU+var(fmOrd2$ETA[[2]]$b)+var(fmOrd2$ETA[[3]]$b)+fmOrd2$varE)
cat(sprintf('h2 = %f\n',h2))
## Lev_17
# step one, create Lev_17
data_476_477_wo_check1 <- data_476_477_wo_check1 %>%
  mutate(Lev_17 = if_else(Hjd_17 > 0, 1, 0)) %>%
  mutate(Lev_17 = ifelse(is.na(Lev_17), 0, Lev_17))
##### check distribution of Lev_17
data_476_477_wo_check1 %>% 
  count(Lev_17)
yBin <- data_476_477_wo_check1$Lev_17
fmBin2<-BGLR(y=yBin,response_type='ordinal',ETA=ETA, nIter=12000, burnIn=2000, saveAt = "Save/saverun_BGLR_H_Lev17")
save(fmBin2,file='Save/fm_H_lev17.rda')
head(fmBin2$probs)
##### h2
h2 <- fmBin2$ETA[[4]]$varU/(fmBin2$ETA[[4]]$varU+var(fmBin2$ETA[[2]]$b)+var(fmBin2$ETA[[3]]$b)+fmBin2$varE)
cat(sprintf('h2 = %f\n',h2))
```

## BLR

```{r}
##### pedigree NRM for comparison with Opsel
y <- data_476_477_wo_check2$Hjd_17
#Trial <- data_476_477_wo_check$Trial
data_476_477_wo_check2 <- data_476_477_wo_check2 %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  mutate(Trial = as.factor(Trial))
X <- model.matrix(~ Trial, data = data_476_477_wo_check2)
Z1 <- model.matrix(~ Trial:Ruta, data_476_477_wo_check2)
Z2 <-  model.matrix(~ Trial:Family_id, data_476_477_wo_check2)

# Fitting the model
### Runs the Gibbs sampler
fm1<-BLR(y=y, XL=X,XF = Z1,XR = Z2,GF=list(ID=1:nrow(Amat2),A=Amat2),
                           nIter=5500,burnIn=500,thin=1, saveAt = "Save/saverun_BLR_A_Hjd17")
save(fm1,file='Save/fm_BLR_A_hjd17.rda')
####### H matrix model
y <- data_476_477_wo_check1$Hjd_17
#Trial <- data_476_477_wo_check$Trial
data_476_477_wo_check1 <- data_476_477_wo_check1 %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  mutate(Trial = as.factor(Trial))
X <- model.matrix(~ Trial, data = data_476_477_wo_check1)
Z1 <- model.matrix(~ Trial:Ruta, data_476_477_wo_check1)
Z2 <- model.matrix(~ Trial:Family_id, data_476_477_wo_check1)

#3# Fitting the model
fm22<-BLR(y=y, XL=X,XF = Z1,XR = Z2,GF=list(ID=1:nrow(Hmat2),A=Hmat2),
                           nIter=12000,burnIn=2000,thin=2, saveAt = "Save/saverun_BLR_G_Hjd17")
save(fm22,file='Save/fm_BLR_H_hjd17.rda')

```

## Additional plots of BGLR analyses

```{r}
############ h2 ##########
# Open the binary file in read mode
con <- file("Save/saverun_BGLR_A_Hjd17ETA_2_b.bin", "rb")

# Read the binary data
#data1 <- readBin(con, numeric())
#data2 <- readBin(con, numeric())
# Initialize the data vector
data1 <- integer()

# Read the binary data in chunks
while (length(chunk <- readBin(con, what = "integer")) > 0) {
  data1 <- c(data1, chunk)
}

# Close the file connection
close(con)


```

## Inference using INLA
Including nested random effects family within trial and plot within trial. The function model.matrix() can be used with a formula that includes the nesting of the variable to create the matrices of the random effects
```{r}
inla.setOption("num.threads", 7)
# Calculate A and H inverse
# Compute the Cholesky decomposition
L <- chol(Amat2)
# Invert the matrix using the Cholesky decomposition
A_inv <- solve(t(L) %*% L)

L <- chol(Hmat2)
H_inv <- solve(t(L) %*% L)


write.table(x = data_476_477_wo_check1,)
data_476_477_wo_check1 <- data_476_477_wo_check1 %>% 
  mutate(ID = seq(1:nrow(data_476_477_wo_check1)))
data_476_477_wo_check2 <- data_476_477_wo_check2 %>% 
  mutate(ID = seq(1:nrow(data_476_477_wo_check2)))

# get the indices and values of the non-zero elements
ind_nnz <- which(A_inv != 0, arr.ind = TRUE)
i <- ind_nnz[, 1]
j <- ind_nnz[, 2]
x <- A_inv[ind_nnz]
# create the sparse matrix
A_inv_sp <- sparseMatrix(i = i, j = j, x = x, dims = dim(A_inv))
ind_nnz <- which(H_inv != 0, arr.ind = TRUE)
i <- ind_nnz[, 1]
j <- ind_nnz[, 2]
x <- H_inv[ind_nnz]
# create the sparse matrix
H_inv_sp <- sparseMatrix(i = i, j = j, x = x, dims = dim(H_inv))

data_476_477_wo_check1 <- data_476_477_wo_check1 %>%
  mutate(Trial = as.factor(Trial)) %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Genotype_id = as.factor(Genotype_id)) # %>%
data_476_477_wo_check2 <- data_476_477_wo_check2 %>%
  mutate(Trial = as.factor(Trial)) %>%
  mutate(Ruta = as.factor(Ruta)) %>%
  mutate(Genotype_id = as.factor(Genotype_id))

# Create nested random effects
# Incidence matrices
#Zi <- as(model.matrix( ~ 0 + Trial, data = data_476_477_wo_check2), "Matrix")
Zij <- as(model.matrix( ~ 0 + Trial:Ruta, data = data_476_477_wo_check2), "Matrix")
#Zijl <- as(model.matrix( ~ 0 + Trial:Ruta:Rad, data = data_upd), "Matrix")
#Zijm <- as(model.matrix( ~ 0 + Trial:Ruta:Planta, data = data_upd), "Matrix")
Zik <- as(model.matrix( ~ 0 + Trial:Family_id, data = data_476_477_wo_check2), "Matrix") # G x E effect
# Before model fitting, indices for the nested random effects must be created
data_476_477_upd <- data_476_477_wo_check2 %>%
  mutate(IDij = seq(1:nrow(data_476_477_wo_check2))) %>%
  mutate(IDik = seq(1:nrow(data_476_477_wo_check2)))

############## Hjd_17
# Fit the model with interaction terms
model = Hjd_17 ~ 1 + Trial + f(ID, model = "generic2", Cmatrix = A_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + # , hyper = prec.prior
        f(IDik, model = "z", Z = Zik)
#Fit INLA

# model = Hjd_17 ~ 1 + Trial + f(ID, model="generic0", Cmatrix=Ainv.sparse) # Ainv_sp
#lm1 <- lm(Hjd_17 ~ 1 + Trial + Trial:Ruta + Trial:Family_id,data = data_upd)
#summary(lm1)

#Y_adj <- resid(lm1)
#data_upd
#IDsave <- names(Y_adj)
#data_upd2 <- data.frame(data_upd[IDsave,],Y_adj)
#ainv.upd2 <- ainverse(data_upd2[,c("Genotype_id","sire","dam")])
#Ainv_sp2 <- sparseMatrix(i = ainv.upd2[,1], j = ainv.upd2[,2], x = ainv.upd2[,3], dims = c(max(ainv.upd2[,1]), max(ainv.upd2[,1])))
#colnames(Ainv_sp2) <- rownames(Ainv_sp2) <- seq(1,nrow(Ainv_sp2))
#Ainv.sparse2 = as(Ainv_sp2, "TsparseMatrix")
#model = Y_adj ~ 1 + f(ID, model = "generic0", Cmatrix = Ainv.sparse2) 

#data_tmp <- data.frame(ID = seq(1,nrow(data_upd2)),Y_adj) #

#rownames(data_tmp) <- seq(1,nrow(data_upd2))

fit_hjd17 = inla(model, data=data_476_477_upd, verbose=TRUE,
           control.compute=list(config=T,dic=T,mlik=T,cpo=T)) # , 
           control.predictor = list(compute = TRUE)
data_476_477_updH <- data_476_477_wo_check1 %>%
  mutate(IDij = seq(1:nrow(data_476_477_wo_check1))) %>%
  mutate(IDik = seq(1:nrow(data_476_477_wo_check1)))
#### H
model = Hjd_17 ~ 1 + Trial + f(ID, model = "generic2", Cmatrix = H_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)
fit_hjd17_H = inla(model, data=data_476_477_updH, verbose=TRUE,
           control.compute=list(config=T,dic=T,mlik=T,cpo=T)) # , 
           control.predictor = list(compute = TRUE)
################# hjd_7

model = Hjd_07 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=A_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)
#+ f(Ruta, by = Trial, model = "iid") + f(Family_id, by = Trial, model = "iid")
fit_hjd7 = inla(model, data=data_476_477_upd, verbose=TRUE,
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))
#### H
model = Hjd_07 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=H_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)

fit_hjd7_H = inla(model, data=data_476_477_updH, verbose=TRUE,
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))

############### Sprant17
model = Sprant_17 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=A_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)
fit_sprant17 = inla(model, data=data_476_477_upd, verbose=TRUE,family = "poisson",
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))
#### H
model = Sprant_17 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=H_inv_sp) + 
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)
fit_sprant17_H = inla(model, data=data_476_477_updH, verbose=TRUE,family = "poisson",
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))
################ Lev_17
model = Lev_17 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=A_inv_sp) +
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)

fit_vit17 = inla(model, data=data_476_477_upd, verbose=TRUE,family = "binomial",
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))
### H
model = Lev_17 ~ 1 + Trial + f(Genotype_id, model="generic2", Cmatrix=H_inv_sp) +
        f(IDij, model = "z", Z = Zij) + 
        f(IDik, model = "z", Z = Zik)

fit_vit17_H = inla(model, data=data_476_477_updH, verbose=TRUE,family = "binomial",
           control.compute=list(config=T,dic=T,mlik=T,cpo=T))


sf <- summary(fit_hjd17); sf
plot(fit_hjd17)
plot(data_upd$Hjd_17, fit_hjd17$summary.fitted.values$mean, main="Fitting result")
plot(fit_hjd17$residuals$deviance.residuals,data_476_477_upd$Hjd_17,xlab = "Residuals",ylab = "Hjd_17")

#Heritability
h2 <- (1/sf[[4]][2,1])/((1/sf[[4]][2,1])+(1/sf[[4]][1,1]) + (1/sf[[4]][4,1]) + (1/sf[[4]][5,1])); h2 # (1/sf[[4]][3,1]) + 


#MSEtest <- sum((fit_hjd17$summary.fitted.values$mean[c(split+1):n] - ytest)^2)/length(ytest)
#print(MSEtest)
#print(cor.test(fit_hjd17$summary.fitted.values$mean[c(split+1):n],ytest, method = 'kendall'))
#print(cor(fit_hjd17$summary.fitted.values$mean[c(split+1):n],ytest))

#Sample from posterior, if we want to calculate 95% CI of MSE and her
fit.samples = inla.posterior.sample(100, fit_hjd17)
hist(fit.samples)
```

# ASREML
## Preparation

```{r}
# Generate a complete set of combinations for Prow and Ppos
complete_combinations <- expand.grid(Prow = 1:30, Ppos = 1:90)

# Find the missing combinations
missing_combinations <- anti_join(complete_combinations, S23F8820476, by = c("Prow", "Ppos"))

# View the missing combinations
head(missing_combinations)
```

## Inference

```{r}
# Univariate runs to get starting values
# Get Ainverse on sparse form
ainv.476_477 <- ainverse(pedigree3[,1:3])
data_476_477_wo_check <- data_476_477_wo_check %>%
  #mutate(Genotype_id = as.factor(Genotype_id)) %>%
  mutate(Family_id = as.factor(Family_id)) %>%
  mutate(Trial = as.factor(Trial)) %>%
  mutate(Prow = as.factor(Prow)) %>%
  mutate(Ppos = as.factor(Ppos))
res.h17 <- asreml(
      fixed = Hjd_17 ~ 1 + Trial,
      random = ~ vm(Genotype_id, ainv.476_477) + 
      idv(Trial):idv(Family_id) +
      idv(Trial):idv(Ruta),
    #  idv(Trial):idv(Ruta):idv(Rad) + 
    #  idv(Trial:idv(Ruta):idv(Planta),
      residual = ~ units, #dsum(~ar1(Ppos):ar1(Prow) | Trial),
 # units,
      data = data_476_477_wo_check,
      na.action = na.method(x = "include"),
      maxit = 1000
    )
summary.asreml(res.h17) 
# calculate the heritability
cat(sprintf('Heritability of trait Hjd_17:\n'))
vpredict(res.h17, h2 ~ V3 / (V1 + V2 + V3 + V4))
res.h7 <- asreml(
      fixed = Hjd_07 ~ 1 + Trial,
      random = ~ vm(Genotype_id, ainv.476_477) + 
      idv(Trial):idv(Family_id) +
      idv(Trial):idv(Ruta),
    #  idv(Trial):idv(Ruta):idv(Rad) + 
    #  idv(Trial:idv(Ruta):idv(Planta),
      residual = ~ units, #dsum(~ar1(Ppos):ar1(Prow) | Trial),
 # units,
      data = data_476_477_wo_check,
      na.action = na.method(x = "include"),
      maxit = 1000
    )
summary.asreml(res.h7) 
# calculate the heritability
cat(sprintf('Heritability of trait Hjd_7:\n'))
vpredict(res.h7, h2 ~ V3 / (V1 + V2 + V3 + V4))
res.htv17 <- asreml(
      fixed = Htv_17 ~ 1 + Trial,
      random = ~ vm(Genotype_id, ainv.476_477) + 
      idv(Trial):idv(Family_id) +
      idv(Trial):idv(Ruta),
    #  idv(Trial):idv(Ruta):idv(Rad) + 
    #  idv(Trial:idv(Ruta):idv(Planta),
      residual = ~ units, #dsum(~ar1(Ppos):ar1(Prow) | Trial),
 # units,
      data = data_476_477_wo_check,
      na.action = na.method(x = "include"),
      maxit = 1000
    )
summary.asreml(res.htv17) 
# calculate the heritability
cat(sprintf('Heritability of trait Htv_17:\n'))
vpredict(res.htv17, h2 ~ V3 / (V1 + V2 + V3 + V4))
res.vit7 <- asreml(
      fixed = Vit_07 ~ 1 + Trial,
      random = ~ vm(Genotype_id, ainv.476_477) + 
      idv(Trial):idv(Family_id) +
      idv(Trial):idv(Ruta),
    #  idv(Trial):idv(Ruta):idv(Rad) + 
    #  idv(Trial:idv(Ruta):idv(Planta),
      residual = ~ units, #dsum(~ar1(Ppos):ar1(Prow) | Trial),
 # units,
      data = data_476_477_wo_check,
      na.action = na.method(x = "include"),
      maxit = 1000
    )
summary.asreml(res.vit7) 
# calculate the heritability
cat(sprintf('Heritability of trait Sprant_17:\n'))
vpredict(res.vit7, h2 ~ V3 / (V1 + V2 + V3 + V4))
res.sprant17 <- asreml(
      fixed = Sprant_17 ~ 1 + Trial,
      random = ~ vm(Genotype_id, ainv.476_477) + 
      idv(Trial):idv(Family_id) +
      idv(Trial):idv(Ruta),
    #  idv(Trial):idv(Ruta):idv(Rad) + 
    #  idv(Trial:idv(Ruta):idv(Planta),
      residual = ~ units, #dsum(~ar1(Ppos):ar1(Prow) | Trial),
 # units,
      data = data_476_477_wo_check,
      na.action = na.method(x = "include"),
      maxit = 1000
    )
summary.asreml(res.sprant17) 
# calculate the heritability
cat(sprintf('Heritability of trait Sprant_17:\n'))
vpredict(res.sprant17, h2 ~ V3 / (V1 + V2 + V3 + V4))
###############################################################
## Multi-trait run
## Get and change starting values
mult.start <- asreml(
      fixed = cbind(Hjd_17, Htv_17, Hjd_07, Sprant_17) ~  trait
        + trait:Trial,
      random = ~ us(trait):vm(Genotype_id, ainv.476_477) + 
        us(trait):idv(Trial):idv(Family_id) +
        us(trait):idv(Trial):idv(Ruta),
      residual = ~ id(units):us(trait),
      data = data_476_477_wo_check,
      na.action = na.method(x = "include", y = "include"),
      maxit = 1000,
      start.values = TRUE
    )
  print(mult.start)$vparameters.table
  # Use grep to find rows with 'Sprant' exactly two times
traits <- c("Hjd_17", "Htv_17", "Hjd_07", "Sprant_17")
lt <- length(traits)
start.v <- rbind(res.h17$vparameters,res.htv17$vparameters,res.h7$vparameters,res.sprant17$vparameters) %>%
  as.data.frame() %>%
  mutate(Trait = traits)
vi <- mult.start$vparameters.table
# Initialize an empty vector to store the indices
combined_indices <- data.frame()
for (i in 1:lt){
  index <- grep(paste(traits[i],".*",traits[i],sep=""),mult.start$vparameters.table[,'Component'])
  vi[index,"Value"] <- t(start.v[i,c(3,2,1,4)])
  # Combine the index vector with the existing ones
  combined_indices <- bind_rows(combined_indices, as.data.frame(index))
}
tot.index <- seq(1:nrow(vi))
# Find elements in 'tot.index' that are not in 'combined_indices'
not_in_combined_indices <- tot.index[!tot.index %in% combined_indices$index]
# set the covariances to zero starting values 
vi[not_in_combined_indices,'Value'] <- 0
##### analyze 
# column major order!
rownames(start.v) <- start.v$Trait
initA <- c(start.v['Hjd_17','vm(Genotype_id, ainv.476_477)'],0,0,0,start.v['Htv_17','vm(Genotype_id, ainv.476_477)'],0,0,start.v['Hjd_07','vm(Genotype_id, ainv.476_477)'],0,start.v['Sprant_17','vm(Genotype_id, ainv.476_477)'])
initF <- c(start.v['Hjd_17','Trial:Family_id!Trial'],0,0,0,start.v['Htv_17','Trial:Family_id!Trial'],0,0,start.v['Hjd_07','Trial:Family_id!Trial'],0,start.v['Sprant_17','Trial:Family_id!Trial'])
initRuta <- c(start.v['Hjd_17','Trial:Ruta!Trial'],0,0,0,start.v['Htv_17','Trial:Ruta!Trial'],0,0,start.v['Hjd_07','Trial:Ruta!Trial'],0,start.v['Sprant_17','Trial:Ruta!Trial'])
initR <- c(start.v['Hjd_17','units!R'],0,0,0,start.v['Htv_17','units!R'],0,0,start.v['Hjd_07','units!R'],0,start.v['Sprant_17','units!R'])
mult.res <- asreml(
      fixed = cbind(Hjd_17, Htv_17, Hjd_07, Sprant_17) ~  trait # 
        + trait:Trial,
      random = ~ us(trait,init = initA):vm(Genotype_id, ainv.476_477) +  # ,init = initA
        us(trait,init = initF):idv(Trial):idv(Family_id) + # ,init = initF
        us(trait,init = initRuta):idv(Trial):idv(Ruta), # ,init = initRuta
      residual = ~ id(units):us(trait, init = initR), # , init = initR
      data = data_476_477_wo_check,
      na.action = na.method(x = "include", y = "include"),
      maxit = 1000,
      workspace=200e06
     # G.param = vi
      #R.param = vi
    )
# ~ dsum(~id(units)|site) as alternative residual different residual VC across trials

# save the data on OPSEL format
# Indiv. Female Male EBV Max
df <- as_tibble(res.h17$coefficients$random)
df$TraitTMP <- rownames(res.h17$coefficients$random)
df$SE <- sqrt(res.h17$vcoeff$random * res.h17$sigma2) 
df <- df %>%
  filter(grepl("Genotype_id", TraitTMP)) %>%
  filter(!grepl("Picea_abies", TraitTMP)) %>%
  rename(EBV = bu)  %>%
# Extract the last part of the string using dplyr and stringr
  mutate(Genotype_id = str_extract(TraitTMP, "(?<=_)[^_]+$")) %>%
  select(-c(TraitTMP))
res.asreml <- merge(df,pedigree3,by="Genotype_id") %>%
  select(-c(SE)) %>%
  select(c(Genotype_id,dam,sire,EBV)) %>%
  mutate(dam = ifelse(grepl('Picea_abies', dam), '0', dam)) %>%
  mutate(sire = ifelse(grepl('Picea_abies', sire), '0', sire)) %>%
  rename(Male = sire, Female = dam, Indiv. = Genotype_id) %>%
  mutate(Max = 1)
write.table(x = res.asreml, file = "Save/EBV_asreml_h17_opsel_input.txt",quote = FALSE, sep = ",",row.names = FALSE)
res.asreml <- read.table(file = "Save/EBV_asreml_h17_opsel_input.txt",sep = ",",header = T)

```



# Notes
Ablation from Machine Learning, removal of components, stability of results